---
title: "Travis Lee"
output: md_document
---
##Loading and preprocessing the data
```{r}
if (!file.exists("./data")) {
        dir.create("./data")     
}
URL <- "https://github.com/TravisLeeTS/RepData_PeerAssessment1/blob/master/activity.zip?raw=true"
path1 <- "./data/activity.zip"
if (!file.exists(path1)) {
        download.file(URL, path1, mode = "wb")
        unzip(path1, exdir = "./data")
}
path2 <- "./data/activity.csv"
activity <- read.csv(path2, na.strings = "NA", stringsAsFactors = FALSE)
activity$date <- as.Date(activity$date)
```

##total, mean and median of the steps taken per day:
```{r}
dailyactivity <- dplyr::summarise(dplyr::group_by(activity, date),
                                  totalsteps = sum(steps),
                                  meansteps = mean(steps, na.rm = TRUE),
                                  mediansteps = median(steps, na.rm = TRUE))
print.data.frame(dailyactivity)
```
##Histogram of the total number of steps taken each day
```{r}
library(ggplot2)
g <- ggplot(data = dailyactivity, aes(date, totalsteps))
gg <- g + geom_bar(stat = "identity") + theme_bw() + xlab("Date") + ylab("Total Steps") + ggtitle("Histogram")
gg
```

##The average daily activity pattern
```{r}
intervalmean <- dplyr::summarise(dplyr::group_by(activity, interval),
                                 intmean = mean(steps, na.rm = TRUE))
g2 <- ggplot(data = intervalmean, aes(interval, intmean))
gg2 <- g2 + geom_line() + theme_bw() + xlab("Interval") + ylab("Average Steps per Day") + ggtitle("Time Series Graph")
gg2
```

##The 5-minute interval on average of all days which contains the maxium number of step
```{r}
intervalmean[which.max(intervalmean$intmean), ]
#At the 835th 5 minutes interval, it reached the max: 206 steps.
```

##Imputing missing values
#Calculating the number of NA's in step variable
```{r}
sum(is.na(activity$steps))
```

##Filling in all missing values using 5 minute-interval means in a new dataset(newactivity)
```{r}
fill <- function(row) {
        intervalmean[intervalmean[, 1] == row[1, 3], 2]
}
newactivity <- activity
for (i in 1:nrow(activity)) {
        if (is.na(activity[i, 1])) {
                newactivity[i, 1] <- fill(activity[i, ])
        }
}
```

##Histogram of the total number of steps taken each day with no NA data
```{r}
newdailyactivity <- dplyr::summarise(dplyr::group_by(newactivity, date),
                                      totalsteps = sum(steps),
                                      meansteps = mean(steps, na.rm = TRUE),
                                      mediansteps = median(steps, na.rm = TRUE))
newg <- ggplot(data = newdailyactivity, aes(date, totalsteps))
newgg <- newg + geom_bar(stat = "identity") + theme_bw() + xlab("Date") + ylab("Total Steps") + ggtitle("Histogram without NA")
newgg
print.data.frame(dailyactivity)
```

##Impact of imputting data on the stimates of the total daily number of steps
```{r}
summary(activity$steps)
summary(newactivity$steps)
#Simialr mean & median but new dataset has higher 3rd quatile
```

##Differences in activity patterns between weekdays and weekends
#Creating a new factor for weekdays and weekends:
```{r}
weekends1 <- c("Saturday", "Sunday")
newactivity$wDay <- factor((weekdays(newactivity$date) %in% weekends1),
                         levels = c(FALSE, TRUE),
                         labels = c("weekday", "weekend"))
```

#Calculating the mean of steps by interval and weekend or weekday:
```{r}
intervalmean2 <- dplyr::summarise(dplyr::group_by(newactivity, interval, wDay),
                                  intmean2 = mean(steps, na.rm = TRUE))
```

#Plotting data:
```{r}
library(lattice)
xyplot(intmean2 ~ interval | wDay, data = intervalmean2, type = "l",
       layout = c(1, 2), xlab = "Interval", ylab = "Average of steps taken")
```